---
title: "COVID-19 India Statewise Vaccine Data Analysis Dashboard"
author: "Satish Dudhat"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(plotly)
library(tidyverse)
library(treemap)
library(knitr)

# Load dataset

#df <- read_csv(file.choose())  
df <- read_csv("C:/Users/hp/Desktop/google learning/Data Anal/R Programming/R-Projects/archive (4)/COVID-19 India Statewise Vaccine Data.csv")

# Data Cleaning & Transformation

df <- df %>% 

replace_na(list(Total_Vaccination_Doses = 0, Dose_1 = 0, Dose_2 = 0, Precaution_18_59 = 0)) %>%
  
mutate(
  Vaccination_Coverage = (Total_Vaccination_Doses / Population) * 100,
  Booster_Coverage = (Precaution_18_59 / Population) * 100,
  Fully_Vaccinated_Percentage = (Dose_2 / Population) * 100,
  Partially_Vaccinated_Percentage = (Dose_1 / Population) * 100)

# Key Statistics

total_vaccination <- sum(df$Total_Vaccination_Doses, na.rm = TRUE)
total_dose1 <- sum(df$Dose_1, na.rm = TRUE)
total_dose2 <- sum(df$Dose_2, na.rm = TRUE)
total_precaution <- sum(df$Precaution_18_59, na.rm = TRUE)

cat("Total Vaccination Doses Administered in India:", total_vaccination, "\n")
cat("Total Dose 1 Administered:", total_dose1, "\n")
cat("Total Dose 2 Administered:", total_dose2, "\n")
cat("Total Precaution Doses (18-59 age group):", total_precaution, "\n")

# Top & Bottom 5 States

top_states <- df %>% arrange(desc(Fully_Vaccinated_Percentage)) %>% head(5)
bottom_states <- df %>% arrange(Fully_Vaccinated_Percentage) %>% head(5)
```

Statewise Vaccination
=======================================================================

Row
-----------------------------------------------------------------------

### Vaccination Doses per State (Scatterplot)

```{r}
p1 <- ggplot(df, aes(x = reorder(State_UTs, Total_Vaccination_Doses), y = Total_Vaccination_Doses, text = paste('State:', State_UTs, '<br>Total Doses:', Total_Vaccination_Doses))) +
  geom_point(color = "blue", size = 2) +
  coord_flip() +
  labs(title = "Total Vaccination Doses vs State", x = "State/UT", y = "Total Vaccination Doses") +
  theme_minimal()

ggplotly(p1, tooltip = "text")

```

### Top 5 States with Highest Vaccination Coverage

```{r}
p2 <- ggplot(top_states, aes(x = reorder(State_UTs, Total_Vaccination_Doses), y = Total_Vaccination_Doses, fill = State_UTs)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 5 States with Highest Vaccinations", x = "State/UTs", y = "Total Vaccination Doses") +
  scale_fill_brewer(palette = "Set3")

ggplotly(p2, tooltip = c("State_UTs", "Total_Vaccination_Doses"))

```
Dose Comparison / Booster Dose
=======================================================================
Row
-----------------------------------------------------------------------

### Dose-1 vs Dose-2 Comparison

```{r}
p3 <- ggplot(df, aes(x = reorder(State_UTs, Dose_1))) +
  geom_bar(aes(y = Dose_1, text = paste('State:', State_UTs, '<br>Dose 1:', Dose_1)), stat = "identity", fill = "blue", alpha = 0.6) +
  geom_bar(aes(y = Dose_2, text = paste('State:', State_UTs, '<br>Dose 2:', Dose_2)), stat = "identity", fill = "orange", alpha = 0.6) +
  coord_flip() +
  labs(title = "Dose-1 vs Dose-2 Comparison Across States", x = "State/UTs", y = "Doses Administered")

ggplotly(p3, tooltip = "text")

```
### Booster Dose Coverage by State

```{r}
p4 <- ggplot(df, aes(x = reorder(State_UTs, Precaution_18_59), y = Precaution_18_59, text = paste('State:', State_UTs, '<br>Booster Coverage:', Precaution_18_59))) +
  geom_bar(stat = "identity", fill = "purple") +
  coord_flip() +
  labs(title = "Booster Dose Coverage by State", x = "State/UTs", y = "Booster Coverage (%)")

ggplotly(p4, tooltip = "text")

```
Age-wise Vaccination & Precaution Dose
=======================================================================

Row
-----------------------------------------------------------------------

### Interactive Stacked Bar Chart for Age-wise Vaccination
```{r}
df_age <- df %>%
  select(State_UTs, Dose_1_15_18, Dose_2_15_18, Dose_1_12_14, Dose_2_12_14) %>%
  pivot_longer(cols = -State_UTs, names_to = "Dose_Type", values_to = "Count")

p5 <- ggplot(df_age, aes(x = State_UTs, y = Count, fill = Dose_Type, text = paste("Dose Type:", Dose_Type, "<br>Count:", Count))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  labs(title = "Age-wise Vaccination Coverage", x = "State/UT", y = "Doses Administered") +
  theme_minimal()
ggplotly(p5, tooltip = "text")
```

### Pie Chart for Precaution Dose Coverage
```{r}
df_precaution <- df %>%
  summarize(Precaution = sum(Precaution_18_59, na.rm = TRUE),
            Other = sum(Total_Vaccination_Doses, na.rm = TRUE) - sum(Precaution_18_59, na.rm = TRUE)) %>%
  pivot_longer(cols = everything(), names_to = "Category", values_to = "Count")

plot_ly(df_precaution, labels = ~Category, values = ~Count, type = "pie",
        textinfo = "label+percent", hoverinfo = "text",
        text = ~paste(Category, "<br>Total Count:", Count),
        marker = list(colors = c("#deb810", "#dbd7c3"))) %>%
  layout(title = "Precaution Dose Coverage")
